---
title: "Data_visualisation"
author: "Jakob Mørup"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#loading Packages
```{r}
library(pacman)
p_load(tidyverse,
       jsonlite,
       sf,
       raster,
       leaflet,
       shiny,
       leaflet.extras,
       ggpattern)

```




#Loading and preprocessing the BBR data, and turning the point data into plotable datatypes using the sf package
```{r}
#Importing the data from the JSON file it is
bbr_cph_data <- fromJSON("outputs/clean_BBR_cph_data.json")
bbr_aarhus_data <- fromJSON("outputs/clean_BBR_aarhus_data.json")



#flattening it to a data frame and changing the name of a column to get rid of danish letters
flatten <- function(json){
  json <- as.data.frame(json[["BygningList"]])
  names(json)[names(json)=="byg026Opførelsesår"] <- "byg026Year"
  return(json)
}


#filtering for buildings used for recreational,educational,habitation,health, light- and service industry out discarding other buildings, refer to BBRkodelister20210322 for the complete list of codes

clean_bbr_df <-function(df, year_cutoff){
  df %>% 
  filter(!is.na(byg404Koordinat)) %>% 
  filter(byg404Koordinat != "POINT(0 0)") %>% 
  filter(byg406Koordinatsystem == "5") %>% 
  filter(!is.na(byg026Year)) %>% 
  filter(byg021BygningensAnvendelse %in% 
           c("101","120","121","122","131","132","140","150","160","190","321","322","324","331","332","333","334","339","421","422","429","431","432","433","439","441","532","533")) %>% 
  distinct(id_lokalId,.keep_all = TRUE) %>% 
  filter(byg026Year >= year_cutoff)
}


bbr_cph_data_flat <- bbr_cph_data %>%
  flatten() %>% 
  clean_bbr_df(year_cutoff = 1900) %>% 
  mutate(city = "Copenhagen")%>% 
  st_as_sf(wkt = "byg404Koordinat", crs = 25832)

bbr_aarhus_data_flat <-bbr_aarhus_data %>%
  flatten() %>% 
  clean_bbr_df(1900) %>% 
  mutate(city = "Aarhus")%>% 
  st_as_sf(wkt = "byg404Koordinat", crs = 25832)



merged_bbr_data_flat <- rbind(bbr_cph_data %>%
  flatten() %>% 
  clean_bbr_df(year_cutoff = 1900) %>% 
  mutate(city = "Copenhagen"), bbr_aarhus_data %>%
  flatten() %>% 
  clean_bbr_df(1900) %>% 
  mutate(city = "Aarhus")) %>% 
  st_as_sf(wkt = "byg404Koordinat", crs = 25832)

```

#loading distric geoJSONs to get mappable district polygons
```{r}

cph_districts <- st_read("data/cph_districts.geojson")
aarhus_districts <- st_read("data/aarhus_districts.geojson")

cph_districts
aarhus_districts
merged_districts <- rbind(aarhus_districts %>% 
                            dplyr::select(navn = prog_distrikt_navn,geometry) %>% 
                            mutate(city = "Aarhus"),
                          cph_districts %>% 
                            dplyr::select(navn,geometry) %>% 
                            mutate(city = "Copenhagen")
                          )
```

# Creating dataframes for intersections of the district polygons and the point data from BBR
```{r}
intersections_aarhus <- st_intersection(aarhus_districts,bbr_aarhus_data_flat) %>% 
  dplyr::select(id = id_lokalId, navn = prog_distrikt_navn,byg021BygningensAnvendelse,byg026Year,byg054AntalEtager,byg406Koordinatsystem,city,geometry)
intersections_cph <-  st_intersection(cph_districts,bbr_cph_data_flat) %>% 
  dplyr::select(id = id_lokalId,navn, byg021BygningensAnvendelse,byg026Year,byg054AntalEtager,byg406Koordinatsystem,city,geometry)

intersections_merged <- rbind(intersections_aarhus, intersections_cph)


```


#Creating choropleth overlays on basemaps from OpenStreetMaps, showing the number of new buildings by districs in a given year or range of years

```{r}
#proj4def = 'EPSG:25832', "+proj=utm +zone=32 +ellps=GRS80 +units=m +no_defs"


#A function that takes a data frame input and a year value (that can be a list/range of values) and returns a subset of the data frame where the year is in the input year(s)
get_year_df <- function(df,year){
temp_df <- df %>% 
  filter(byg026Year %in% year)
    return(temp_df)
}

#A function that adds a new column to the merged districts with counts of new buildings in each districts in that/those year(s) given a range of years, from the intersections data frame and returns the new dataframe
n_new_buildings <- function(year){
  n_new <- rep(0,length(merged_districts$navn))
  intersections_year <- get_year_df(intersections_merged, year)
  for (i in 1:length(merged_districts$navn)){
    n_new[i] <- nrow(filter(intersections_year, navn == merged_districts$navn[i]))
  }

  temp <- cbind(merged_districts, n_new)
  return(temp)
}

#Function that takes a year range and creates a choropleth map from it, based on BBR data and district geometry
draw_choropleth <- function(year){
temp_df_choropleth <- n_new_buildings(year = year)


#creating a color palette
pal <- colorNumeric("Greens", domain = NULL)

#creating labels for the choropleth map
labels <- sprintf(
  "<strong>%s</strong><br/>%g Nye bygninger</sup>",
  temp_df_choropleth$navn, temp_df_choropleth$n_new
) %>% lapply(htmltools::HTML)


#drawing the choropleth map on top of OpenStreetMaps base map
choro <- leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = st_transform(temp_df_choropleth,4326),
                      fillColor = pal(temp_df_choropleth$n_new),
                      weight = 1,
                      opacity = 1,
                      color = "white",
                      dashArray = "3",
                      fillOpacity = 0.7,
                      highlightOptions = highlightOptions(
                        weight = 5,
                        color = "#666",
                        dashArray = "",
                        fillOpacity = 0.7,
                        bringToFront = TRUE),
                      label = labels,
                      labelOptions = labelOptions(
                        style = list("font-weight" = "normal", padding = "3px 8px"),
                        textsize = "15px",
                        direction = "auto")) %>% 
  addLegend(
    data = temp_df_choropleth,
    position = "bottomright",
    pal = pal,
    values = ~ n_new,
    title = "Antal nye bygninger",
    opacity = 0.5)


return(choro)
}

draw_choropleth(c(1990:1991))

```



## making Histograms of average building height measured in floors.

```{r}
plot_hist <- function(Year, df){
temp_df_histogram <- 
  get_year_df(df, Year) %>%
  dplyr::select(byg026Year, byg054AntalEtager) %>% 
  group_by(byg026Year) %>% 
  summarise_at(vars(byg054AntalEtager),funs(mean(., na.rm =TRUE)))
  

  ggplot(temp_df_histogram,aes(x = byg026Year, y= byg054AntalEtager))+
    geom_col(fill = "#76EE00", color = "black")+
    labs(x= "Years", y= "Mean heigth in floors")+
    ggtitle(paste(df$city[1]))+
    theme_minimal()
}

plot_hist(c(1990:1999), intersections_aarhus)

```

